Fixing Vulnerabilities Potentially Hinders Maintainability

### 2020-05-29 paper/icsme20.tex
fixes

Novel tools have been built to automatically detect software vulnerabilities (e.g., FindBugs, Infer and more). 
Despite these efforts, improving software security is not a trivial task and requires 
implementing patches that might affect software maintainability. 
Our suspicion is that some of these patches may have a negative 
impact on the software maintainability and, possibly, even be the cause of the 
introduction of new vulnerabilities---harming software reliability and introducing technical debt. Therefore, 
in this paper, we present an empirical study on the impact of patches of 
vulnerabilities on software maintenance across open-source software.
%
% Software maintainability can also be defined as 
% the property of software that reveals how easily the software system can be 
% maintained~\cite{IEEEComputerSociety:2014:GSE:2616205} or recover after a failure. 
% Hedegus et al.
% ($2018$)~\cite{HEGEDUS2018313}, analyzed the difference of maintainability between
% groups of refactored elements and non-refactored and conclude that the source
% code elements subjected to refactorings had significantly lower maintainability
% than the elements not modified.
% There is still little information about the impact of patching vulnerabilities
% in open-source software maintainability. \textit{Are developers affecting software
% maintainability when trying to improve software security?}, it is what we want to
% investigate.
%
As ISO does not provide any specific guidelines/formulas to calculate 
maintainability, we resort to Software Improvement Group (SIG)'s web-based source 
code analysis service Better Code Hub (BCH) to compute the software compliance 
with a set of $10$ guidelines/metrics to produce quality software 
based in ISO/IEC $25010$~\cite{Visser:2016:OREILLY}. 
There are other well-known standards and models that have been proposed
to increase software security: Common Criteria~\cite{common:2009} which received
negative criticism regarding the costs associated and poor technical evaluation;
the OWASP Application Security Verification Criteria~\cite{oswap:2009} which is
focused only on web applications, and a model proposed by Xu et al. ($2013$)
~\cite{6616351} for rating software security (arguably, it was one of the
first steps taken by SIG to introduce security on their maintainability model). 
Nevertheless, our study uses BCH to provide an assessment of maintainability in 
software for the following reasons: BCH integrates a broad number of code metrics; and, 
code metrics were empirically validated in previous work~\cite{Bijlsma:2012:FIR:2317098.2317124, 8530041, cruz2019energyoriented}.

@@

Many studies have investigated the relationship between refactorings and
software quality. Previous work focused on object-oriented metrics has evaluated the
impact of refactorings and exhibited proof that quantifying effectively the
impact of refactorings on maintainability may help to choose the appropriate
refactoring type~\cite{1167822}. In contrast to this work, Hegedus et
al.~\cite{HEGEDUS2018313} did not select particular metrics to assess the effect
of refactorings. Instead, statistical tests were used to find the metrics that
have the potential to change significantly after refactorings. Studying the evolution 
of maintainability issues during the development of Android apps, Malavolta et al. ($2018$)~\cite{8530041}
discovered that maintainability decreases over time. Palomba et al.
($2018$)~\cite{Palomba:2018:DIM:3231288.3231337} exhibits proof that code smells
should be carefully monitored by programmers since there is a high correlation
between maintainability aspects and proneness to changes/faults. In 2019, Cruz et 
al.~\cite{cruz2019energyoriented} proposed a formula to calculate maintainability 
based on the BCH's guidelines and measured the impact of energy-oriented fixes 
on software maintainability. The present work uses the same model (BCH) and focuses 
solely on evaluating the impact of security patches on software maintainability.

>>>>

Novel tools have been built to automatically detect software vulnerabilities (e.g., FindBugs, Infer and more). 
Despite these efforts, improving software security is not a trivial task and requires 
implementing patches that might affect software maintainability. 
Our suspicion is that some of these patches may have a negative 
impact on the software maintainability and, possibly, even be the cause of the 
introduction of new vulnerabilities---harming software reliability and introducing technical debt. Therefore, 
in this paper, we present an empirical study on the impact of patches of 
vulnerabilities on software maintenance across open-source software.
%
% Software maintainability can also be defined as 
% the property of software that reveals how easily the software system can be 
% maintained~\cite{IEEEComputerSociety:2014:GSE:2616205} or recover after a failure. 
% Hedegus et al.
% ($2018$)~\cite{HEGEDUS2018313}, analyzed the difference of maintainability between
% groups of refactored elements and non-refactored and conclude that the source
% code elements subjected to refactorings had significantly lower maintainability
% than the elements not modified.
% There is still little information about the impact of patching vulnerabilities
% in open-source software maintainability. \textit{Are developers affecting software
% maintainability when trying to improve software security?}, it is what we want to
% investigate.
%
As ISO does not provide any specific guidelines/formulas to calculate 
maintainability, we resort to Software Improvement Group (SIG)'s web-based source 
code analysis service Better Code Hub (BCH) to compute the software compliance 
with a set of $10$ guidelines/metrics to produce quality software 
based in ISO/IEC $25010$~\cite{Visser:2016:OREILLY}. 
There are other well-known standards and models that have been proposed
to increase software security: Common Criteria~\cite{common:2009} which received
negative criticism regarding the costs associated and poor technical evaluation;
the OWASP Application Security Verification Criteria~\cite{oswap:2009} which is
focused only on web applications, and a model proposed by Xu et al. ($2013$)
~\cite{6616351} for rating software security (arguably, it was one of the
first steps taken by SIG to introduce security on their maintainability model). 
Nevertheless, our study uses BCH to provide an assessment of maintainability in 
software for the following reasons: BCH integrates a broad number of code metrics; and, 
code metrics were empirically validated in previous work~\cite{Bijlsma:2012:FIR:2317098.2317124, 8530041, cruz2019energyoriented, 8785997}.

@@

Many studies have investigated the relationship between refactorings and
software quality. Previous work focused on object-oriented metrics has evaluated the
impact of refactorings and exhibited proof that quantifying effectively the
impact of refactorings on maintainability may help to choose the appropriate
refactoring type~\cite{1167822}. In contrast to this work, Hegedus et
al.~\cite{HEGEDUS2018313} did not select particular metrics to assess the effect
of refactorings. Instead, statistical tests were used to find the metrics that
have the potential to change significantly after refactorings. Studying the evolution 
of maintainability issues during the development of Android apps, Malavolta et al. ($2018$)~\cite{8530041}
discovered that maintainability decreases over time. Palomba et al.
($2018$)~\cite{Palomba:2018:DIM:3231288.3231337} exhibits proof that code smells
should be carefully monitored by programmers since there is a high correlation
between maintainability aspects and proneness to changes/faults. In 2019, Cruz et 
al.~\cite{cruz2019energyoriented} proposed a formula to calculate maintainability 
based on the BCH's guidelines and measured the impact of energy-oriented fixes 
on software maintainability. Recent work,
proposed a new maintainability model to 
measure fine-grained code changes by adapting/extending the BCH model~\cite{8785997}.
The present work uses the same model (BCH) and focuses 
solely on evaluating the impact of security patches on software maintainability.
